---
title: "042623_MOFA2_RNA_PROTEIN_PHENOTYPE"
author: "Amir Chaman Baz"
date: "4/26/2024"
output: html_document
---
#-------------------------------------------------------------------------------
#-----Model 2:This Script runs MOFA2 analysis on RNA, protein and phenotype data 
#-----using a subset of the data to have 50% of all analyzed samples have 
#-----RNA-data and maximize the amount of protein data within.
#-----Outputs correlations data: rho, BH adjusted p-values for each factor and 
#-----associated outcome
#-----Outputs: Figure 1, 2 and 3b of Model 2 chapter in paper
#-------------------------------------------------------------------------------

```{r}
#Select whether to run new MOFA2 analysis or load Mofaobject (Default=2)
cat("run new MOFA model(1) or load from file(2)?: ")
optionp<- readline()

if(!optionp==1){
  optionp<-2
}
```

 
#Load libraries
```{r}
library(MOFA2)
library(readxl)
library(dplyr)
library(purrr)
library(ggplot2)
library(reshape2)
library(ggplotify)
library(gridExtra)
library(gplots)
library(corrplot)
library(MASS)
library(VennDiagram)
library(ggcorrplot)
library(viridis)
library(ggpmisc)
library(sysfonts)
library(showtext)
library(ggpubr)
library(gprofiler2)
library(mixOmics)
library(psych)
library(gtable)
```

```{r}
#-----Add fonts for plots
font.add("baskerville",
         regular = "Libre_Baskerville(1)/LibreBaskerville-Regular.ttf",
         bold = "Libre_Baskerville(1)/LibreBaskerville-Bold.ttf",
         italic = "Libre_Baskerville(1)/LibreBaskerville-Italic.ttf")
showtext.auto()
```

```{r}
#-----Import functions used for plot generation
source("importfunctions.R")

```


```{r}
#-----Load genenames database (EnsDb.Hsapiens.v86.sqlite)
library(RSQLite)
dbfile <- "EnsDb.Hsapiens.v86.sqlite"
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = dbfile)
ensembltables <- dbListTables(con)

ensembldf <- vector("list", length = length(ensembltables))

for (i in seq(along = ensembltables)) {
  ensembldf[[i]] <- dbGetQuery(
    conn = con,
    statement = paste("SELECT * FROM '",
                      ensembltables[[i]],
                      "'",
                      sep = "")
  )
}

genenames <- data.frame(gene_id = ensembldf[[4]]$gene_id,
                        gene_name = ensembldf[[4]]$gene_name)
```

```{r}
#-----Load proteinnames from database (UniprotKB)
file <- "HUMAN_9606_idmapping_selected.tab/HUMAN_9606_idmapping_selected.tab"
proteindf <- read.csv(file,
                      sep = "\t",
                      header = FALSE)

proteinnames <- data.frame(protein.id = proteindf$V1,
                           protein.name = proteindf$V2)

file <- NULL

#Remove substring "_HUMAN" from proteinnames
proteinnames$protein.name <-substring(proteinnames$protein.name,
                                      1,
                                      nchar(proteinnames$protein.name) - 6)

```

 
#Load data from external files
```{r}
#Loading data from external files
load(file.choose())
pdf$PatientID <- gsub("_V.", "", rownames(pdf))

load(file.choose())
rna_counts <- data.frame(v$E)

df <- read_excel(file.choose(), sheet = 1)
df <- as.data.frame(df)
names(df)[names(df) == "PatientID...1"] <- "sample"
```

```{r}
#import Gprofiler results (RDS)
if(optionp==2){
gsearesults <- readRDS(file.choose()) 
} 
```
  

```{r}
#Calculate StroopInterference and TMT 
stroop2V2 <-
  ((60 - df$StroopCardIIErrorsV2) / 60) / df$StroopCardIITimeV2
stroop2V4 <-
  ((60 - df$StroopCardIIErrorsV4) / 60) / df$StroopCardIITimeV4
stroop3V2 <-
  ((60 - df$StroopCardIIIErrorsV2) / 60) / df$StroopCardIIITimeV2
stroop3V4 <-
  ((60 - df$StroopCardIIIErrorsV4) / 60) / df$StroopCardIIITimeV4
df$StroopInterferenceV2 <- stroop3V2 / stroop2V2
df$StroopInterferenceV4 <- stroop3V4 / stroop2V4

df$TMTV2 <- df$TMTBV2 / df$TMTAV2
df$TMTV4 <- df$TMTBV4 / df$TMTAV4


```


```{r}
#Selection of clinical outcomes from list
#seperate two timepoints and merge 

#select V2 vars only
clindf<- df[colnames(df)[grep("*\\V2$", colnames(df))]] 
#Select V4 vars only
clindfv4<- df[colnames(df)[grep("*\\V4$", colnames(df))]]


#remove all variables starting with Stroop
clindf<- clindf[-grep("^Stroop.*$", colnames(clindf))] 
clindfv4<- clindfv4[-grep("^Stroop.*$", colnames(clindfv4))] 
      
#Add Stroopinterference and TMT scores
clindf$StroopInterferenceV2<- df$StroopInterferenceV2
clindf$TMTV2<- df$TMTV2

clindfv4$StroopInterferenceV4<- df$StroopInterferenceV4
clindfv4$TMTV4<- df$TMTV4
      
#Remove TMTA en TMTB, PreBorg, MIRS, MCgillpain, l5ENMO, M5ENMO, SIPScore
omitlist<- c(paste0("TMTBV2|TMTAV2|PreBORGV2|MIRSV2|McGillPainV2|",
                          "L5ENMOV2|M5ENMOV2|SIPScoreV2"))
clindf<- clindf[!grepl(omitlist, colnames(clindf))]
      
omitlist<- c(paste0("TMTBV4|TMTAV4|PreBORGV4|MIRSV4|McGillPainV4|",
                          "L5ENMOV4|M5ENMOV4|SIPScoreV4"))
clindfv4<- clindfv4[!grepl(omitlist, colnames(clindfv4))]

#Add samplenames to both df      
samplenames<- substring(df$sample, 1, nchar(df$sample)-1)
rownames(clindf)<- paste0(samplenames,"_V2")
rownames(clindfv4)<- paste0(samplenames,"_V4")
      
```

```{r}
#Remove V2/V4 from colnames, merge df's
colnames(clindfv4)<- substring(colnames(clindfv4),
                                     1, nchar(colnames(clindfv4))-2)
colnames(clindf)<- substring(colnames(clindf), 
                                     1, nchar(colnames(clindf))-2)
clinicaloutcomescoresdf<- rbind(clindf, clindfv4)
```

```{r}
#Filtered phenotypes  (Apply IQR_filter to clinicaloutcomes)
clinfiltered<- lapply(clinicaloutcomescoresdf, function(x)IQR_filter(x,3)) 
clinfiltered<- as.data.frame(clinfiltered)
rownames(clinfiltered)<- rownames(clinicaloutcomescoresdf)

numfiltered<-sum(is.na(clinfiltered)) - sum(is.na(clinicaloutcomescoresdf))
paste0(numfiltered, " outliers filtered ") #Should be 5 outliers instead of 10 since we omitted clinical outcomes above

clinicaloutcomescoresdf<- clinfiltered


```


```{r}
#Make new dataframe clinoutcomes with transformed variables
#based on BoxCox Transformation
clinoutcomes<-clinicaloutcomescoresdf

clinoutcomes$CISFatigue<- clinicaloutcomescoresdf$CISFatigue^2 
clinoutcomes$BDIFs<- sqrt(clinicaloutcomescoresdf$BDIFs)
clinoutcomes$SSLNScore<- 1/sqrt(clinicaloutcomescoresdf$SSLNScore)
clinoutcomes$SSLDScore<- 1/sqrt(clinicaloutcomescoresdf$SSLDScore)
clinoutcomes$JFCS<- 1/sqrt(clinicaloutcomescoresdf$JFCS)
clinoutcomes$CISactivity<- clinicaloutcomescoresdf$CISactivity^2


```


```{r}
#Z-score scaling for each clinical outcome measurement by variable
clinoutcomes_scaled<- as.data.frame(lapply(clinoutcomes, 
                                    function(x) 
                                            scale(x, 
                                                  center=TRUE, 
                                                  scale = TRUE)))
rownames(clinoutcomes_scaled)<- rownames(clinoutcomes)


```

```{r}
#Z-score scaling for each rpc protein 
rpc_scaled<- as.data.frame(apply(rpc, 1, 
                                  function(x) 
                                      scale(x, 
                                            center=TRUE, 
                                            scale = TRUE)))


rpc_scaled<-t(rpc_scaled)
colnames(rpc_scaled)<- colnames(rpc)
rpc_scaled<- as.data.frame(rpc_scaled)
```


```{r}
#Z-score scaling for each rna
rna_scaled<- as.data.frame(apply(rna_counts, 1, 
                                  function(x) 
                                      scale(x, 
                                            center=TRUE, 
                                            scale = TRUE)))


rna_scaled<-t(rna_scaled)
colnames(rna_scaled)<- colnames(rna_counts)
rna_scaled<- as.data.frame(rna_scaled)



```


   
```{r}
#-----Inverse clinical outcomes so they have have the same directionality (higher is better)
clinoutcomes_scaled$MDHI <- clinoutcomes_scaled$MDHI*(-1)
clinoutcomes_scaled$CISFatigue <- clinoutcomes_scaled$CISFatigue*(-1)
clinoutcomes_scaled$INQOLQolScore <- clinoutcomes_scaled$INQOLQolScore*(-1)
clinoutcomes_scaled$BDIFs <- clinoutcomes_scaled$BDIFs*(-1)
clinoutcomes_scaled$AEScScore <- clinoutcomes_scaled$AEScScore*(-1)
clinoutcomes_scaled$TMT <- clinoutcomes_scaled$TMT*(-1)
clinoutcomes_scaled$ASBQ <- clinoutcomes_scaled$ASBQ*(-1)
clinoutcomes_scaled$SSLDScore <- clinoutcomes_scaled$SSLDScore*(-1)
clinoutcomes_scaled$SSLNScore <- clinoutcomes_scaled$SSLNScore*(-1)
clinoutcomes_scaled$JFCS <- clinoutcomes_scaled$JFCS*(-1)
clinoutcomes_scaled$IMQ <- clinoutcomes_scaled$IMQ*(-1)
clinoutcomes_scaled$CISactivity <- clinoutcomes_scaled$CISactivity*(-1)
clinoutcomes_scaled$FDSS <- clinoutcomes_scaled$FDSS*(-1)

#CSI and AESI lower is better, added this
clinoutcomes_scaled$CSI <- clinoutcomes_scaled$CSI*(-1)
clinoutcomes_scaled$AESI <- clinoutcomes_scaled$AESI*(-1)

```


```{r}
rna_counts2<- rna_scaled
rpc2<- rpc_scaled
#c----If you want to run the analysis on 1 timepoint:
#c----rpc2<-rpc_scaled for scaled by protein
#c----Select V2 only:
#c----rna_counts2<- rna_counts2 %>% select(-ends_with("V4"))
#c----rpc2<- rpc2 %>% select(-ends_with("V4"))
#c----head(rna_counts2)
#c----head(rpc2)

#c----Add sample names and transpose clinical data to match rna and rpc
#c----rownames(clinicaloutcomescoresdf)<-paste0(substring(df$sample,
#c----                                                    1, 
#c----                                                    nchar(df$sample)-1),
#c----                                                    "_V2")
      
clinoutcomes_scaled<- t(clinoutcomes_scaled)
clinoutcomes_scaled<- as.data.frame(clinoutcomes_scaled)
```

```{r}
#Select 100 samples, subset contains all samples with RNA (n=54), 
#maximize n samples who have protein data
#Sample all RNA, all protein, all DM1Activ, All SMWT, 
#<20% missing clinical outcome, then 
#stratify male/female/timepoints


nneeded<- 100- length(colnames(rna_counts2)) #46 samples to select

mask<- !colnames(clinoutcomes_scaled) %in% colnames(rna_counts2)
clin_no_rna<-clinoutcomes_scaled[,mask]

mask<- !colnames(rpc2) %in% colnames(rna_counts2)
prot_no_rna<- rpc2[,mask]

prot_in_clin<- clin_no_rna[, colnames(prot_no_rna)]

#-----Select only samples that have DM1ActivC and SMWT scores
mask<- which(rownames(prot_in_clin)=="DM1ActivC"|rownames(prot_in_clin)=="SMWT")
DM1SMWT<- prot_in_clin[mask, ]

mask<-unlist(lapply(DM1SMWT, function(x) !is.na(sum(x))))
prot_in_clin<-prot_in_clin[, mask]

#-----Select only samples with >80% clinical outcomes
mask<- unlist(lapply(prot_in_clin, function(x)(sum(!is.na(x))/length(x))>0.8))
prot_in_clin<-prot_in_clin[, mask]


prot_in_clin<- t(prot_in_clin)
prot_in_clin<- as.data.frame(prot_in_clin)
samples<- substring(rownames(prot_in_clin), 1, nchar(rownames(prot_in_clin))-3)



#-----Add Sex data
prot_in_clin$sex<-""
for( y in samples){
    prot_in_clin$sex[which(samples==y)]<- df$SexCode[which(substring(df$sample,1, nchar(df$sample)-1)==y)]   
      }
#-----Add timepoints
prot_in_clin$timepoint<- substring(rownames(prot_in_clin), 
                                         nchar(rownames(prot_in_clin))-1, 
                                         nchar(rownames(prot_in_clin)))      
      
prot_in_clin$samplename<- rownames(prot_in_clin)

#-----Select samples in the subset that have protein data for both timepoints
samplenames<- substring(rownames(prot_in_clin), 1, nchar(rownames(prot_in_clin))-3)
samplecounts<-table(samplenames)
completeprotein<-paste0(names(samplecounts[samplecounts>1]), "_V2") 
      
      
prot_in_clin_V2_subset<- prot_in_clin[grep(pattern = "_V2", rownames(prot_in_clin)),]
prot_in_clin_V2_subset<- prot_in_clin_V2_subset[rownames(prot_in_clin_V2_subset) %in% completeprotein,]
      
set.seed(123)
strat_norna<- prot_in_clin_V2_subset %>% 
group_by(sex, timepoint) %>%
sample_n(size=12)
      
paste0("sex")
table(strat_norna$sex)
paste0("timepoints")
table(strat_norna$timepoint)
#Find sample pair(other timepoint) and add to selection
V4equivalent<- paste0(substring(strat_norna$samplename, 1, nchar(strat_norna$samplename)-3), "_V4")

samplesselected<- c(colnames(rna_counts2), strat_norna$samplename, V4equivalent)
print(samplesselected)

```

```{r}
#-----subset Data based on the sampling above
      
rpc2<- rpc2[,samplesselected[samplesselected %in% colnames(rpc2)]]
rna_counts2<- rna_counts2[,samplesselected[samplesselected %in% colnames(rna_counts2)]]
clinoutcomes_scaled<- clinoutcomes_scaled[,samplesselected[samplesselected %in% colnames(clinoutcomes_scaled)]]
      
      
#-----backup
rna_counts3<- rna_counts2
clinoutcomes_scaled3<- clinoutcomes_scaled

```

```{r}
#-----Make rpc and rna_counts have equal samples by adding NA columns for the 
#-----missing samples in each (This is required by MOFA2 )
      
mask<- !(colnames(clinoutcomes_scaled) %in% colnames(rpc2))
missingsamples<- colnames(clinoutcomes_scaled)[mask]
for (y in missingsamples ){
            rpc2[[y]]<- NA
          }
      
mask<- !(colnames(clinoutcomes_scaled) %in% colnames(rna_counts2))
missingsamples<- colnames(clinoutcomes_scaled)[mask]
for (y in missingsamples ){
     rna_counts2[[y]]<- NA
          }
      
rna_counts2<- rna_counts2[,order(colnames(rna_counts2))]
rpc2<- rpc2[,order(colnames(rpc2))]
clinoutcomes_scaled<- clinoutcomes_scaled[, order(colnames(clinoutcomes_scaled))]
```
 
```{r}
#Build MOFA Object and Set Options
#Default seed
mofadata<-list(rna = as.matrix(rna_counts2), 
                     protein= as.matrix(rpc2), 
                     phenotype=as.matrix(clinoutcomes_scaled))
MOFAobject<- create_mofa(mofadata)
      
      
data_opts<- get_default_data_options(MOFAobject)
model_opts<- get_default_model_options(MOFAobject)
model_opts$num_factors<- 30 #Number of factors
train_opts<- get_default_training_options(MOFAobject)
train_opts$convergence_mode <- "slow"
train_opts$maxiter<- 10000
      
MOFAobject <- prepare_mofa(MOFAobject, 
                                 data_options= data_opts, 
                                 model_options = model_opts, 
                                 training_options = train_opts)

```

#load model or run mofa:
```{r}
mofaobjectfile<- file.choose()
if(optionp==2){
      MOFAobject_ran <- load_model(mofaobjectfile)}else if(optionp==1){
        MOFAobject_ran<- run_mofa(MOFAobject, 
                           outfile=file.choose(), 
                           use_basilisk = FALSE)
                          }
```
 

#load metadata in MOFAobject_ran: 
```{r}
df2<-df
df2$sample<-paste0(substring(df2$sample, 1, nchar(df2$sample)-1))


df3<-df2[1,]
df3$timepoint<- ""
df3<-df3[0,]

#----For each sample V2 and V4, double metadata so every sample has 
#----its own metadata entry
for(g in df2$sample){
          df3<-rbind(df3, 
                     cbind(sample=paste0(g, "_V2"), 
                           df2[which(df2$sample==g),-1], 
                           timepoint=0), 
                     cbind(sample=paste0(g, "_V4"), 
                           df2[which(df2$sample==g),-1], 
                           timepoint=1))
          }


df3$CTGDiagnostic<- as.numeric(df3$CTGDiagnostic)
df3$V2Mode<- as.numeric(df3$V2Mode)
df3$LogCTG_s<- scale(log(df3$CTGDiagnostic), 
                           center=TRUE, 
                           scale=TRUE)
df3$AgeAtOnset<- as.numeric(df3$AgeAtOnset)

#Select only the metadata of the 102 samples:
df3_subset<- df3[df3$sample %in% samplesselected,]  
      
metadatadf<- cbind(df3_subset[!colnames(df3_subset) %in% rownames(clinoutcomes_scaled)],
                                                t(clinoutcomes_scaled))
samples_metadata(MOFAobject_ran) <- metadatadf
      
      
      
metadatadf<- t(metadatadf)
metadatadf<- as.data.frame(metadatadf) 
colnames(metadatadf)<- metadatadf[which(rownames(metadatadf)=="sample"),] 
      
   
      
      
```


```{r}
#manual variance data:
r2<- calculate_variance_explained(MOFAobject_ran)
r2perfactor<- as.matrix(r2$r2_per_factor$group1)
r2perfactor<- melt(r2perfactor)
```



```{r}
#Generate R2 and  p-values of LM heatmap covariates ~ factors
listn <-Corr_factors_R2(MOFAobject_ran, 
                              covariates = c(#"TreatmentCode", 
                                              "V2Age",
                                              "AgeAtOnset",
                                              "SexCode", 
                                              "timepoint",
                                              "CTGDiagnostic"), 
                                              plot = "r")
Z<- listn[[1]] #Retrieve Factor values per sample
covariates <- listn[[2]] #Get covariates
Z<-as.data.frame(Z)


lmheatmap <- covariates[0, ] #Empty df with Covariate names as colnames
lmheatmap_pval <- covariates[0, ]

for(j in 1:30) {
  l <- list()
  pval <- list()
  for (i in colnames(lmheatmap)) {
    testdf <- data.frame(Z = Z[[j]], cov = covariates[[i]])
    
    lmmodel <- summary(lm(cov ~ Z, data = testdf))
    l <- c(l, lmmodel$r.squared)
    pval <- c(pval, lmmodel$coefficients[2, 4])
  }
  
  l <- as.data.frame(t(l))
  colnames(l) <- colnames(lmheatmap)
  lmheatmap <- rbind(lmheatmap, l)
  
  pval <- as.data.frame(t(pval))
  colnames(pval) <- colnames(lmheatmap_pval)
  lmheatmap_pval <- rbind(lmheatmap_pval, pval)
  
}

lmheatmap<- as.data.frame(lmheatmap)
lmheatmap$V2Age<- as.numeric(lmheatmap$V2Age)
lmheatmap$AgeAtOnset<- as.numeric(lmheatmap$AgeAtOnset)
lmheatmap$SexCode<- as.numeric(lmheatmap$SexCode)
lmheatmap$timepoint<- as.numeric(lmheatmap$timepoint)
lmheatmap$CTGDiagnostic<- as.numeric(lmheatmap$CTGDiagnostic)

rownames(lmheatmap)<- r2perfactor$Var1[1:30]
lmheatmap$factor<- r2perfactor$Var1[1:30]
lmmelt<- melt(lmheatmap, id.vars = "factor")

lmheatmap_pval<- as.data.frame(lmheatmap_pval)
lmheatmap_pval$V2Age<- as.numeric(lmheatmap_pval$V2Age)
lmheatmap_pval$AgeAtOnset<- as.numeric(lmheatmap_pval$AgeAtOnset)
lmheatmap_pval$SexCode<- as.numeric(lmheatmap_pval$SexCode)
lmheatmap_pval$timepoint<- as.numeric(lmheatmap_pval$timepoint)
lmheatmap_pval$CTGDiagnostic<- as.numeric(lmheatmap_pval$CTGDiagnostic)

rownames(lmheatmap_pval)<- r2perfactor$Var1[1:30]
lmheatmap_pval$factor<- r2perfactor$Var1[1:30]

lmheatmap_pval<- melt(lmheatmap_pval, value.name = "pval")
lmheatmap_pval<- as.data.frame(lmheatmap_pval)
lmheatmap_pval$pval_adjusted<- p.adjust(lmheatmap_pval$pval, method = "BH")
```



```{r}
#Get all weights (protein, phenotype)
weights<- get_weights(MOFAobject_ran, views = "all", factors = "all")
protein.weights<- as.data.frame(weights$protein)
rna.weights<-as.data.frame(weights$rna)
phenotype.weights<- as.data.frame(weights$phenotype)
```

```{r}
#Visualise overview of Data in MOFAobject 
plot_data_overview(MOFAobject)
```

#Sanity check; verify whether factors are largely uncorrelated:
```{r}
#Sanity check; verify whether factors are largely uncorrelated:
plot_factor_cor(MOFAobject_ran)
```

```{r}
#Get data on correlation of clinical outcomes with factors
corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "r", return_data = TRUE)
corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "pval", return_data = TRUE)
```

#generation of spearman correlations (Return data) 
```{r}
#Factor 6 correlations
for(i in c("IMQ", "ICQ", "JFCS", "SSLIScore", "SSLDScore", "ASBQ", "BDIFs", "AEScScore")){
corrfactortable<-regressor_rna(i, nfactor = 6, nfeature = 25, colby="timepoint", returncorr = TRUE)
print(corrfactortable)
print(paste0(i,":" , "BH adjusted pval:"))
print(corrfactortable$p.adj)
cat("-------------------------------\n\n")
corrfactortable<- NULL
}
```

#-------------------------------------------------------------------------------
#All figures below: 
```{r}
#Scree plot
screeplottable<- r2$r2_per_factor$group1
screeplottable<- as.data.frame(screeplottable)
screeplottable$total<- screeplottable$rna +
    screeplottable$protein + screeplottable$phenotype

screeplotfigure<- ggplot(screeplottable, aes(x=1:30, y= total)) + 
    geom_line(lwd=1) + 
    xlab("Factor") + 
    ylab("Variance explained (%)") + 
    scale_x_continuous(breaks = seq(1,30,1)) +
    scale_y_continuous(breaks=seq(0,30,5))+
    theme_minimal() +
    theme(axis.text.x = element_text(size=15), axis.text.y = element_text(size=15), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(text=element_text(family = "baskerville")) +
  geom_hline(yintercept = 0)+
   geom_vline(xintercept = 0)
		 
print(screeplotfigure)
``` 

```{r}
#Explained variance per factor (Fig1a)

fig1a<- plot_variance_explained(MOFAobject_ran, plot_total = F)+
  theme(text=element_text(family = "baskerville"))+
  ggtitle("(a)")

#Explained variance per covariate (Fig2b)
fig1b<- ggplot(lmmelt, 
       aes(x = variable, y = factor, fill = value)) +
    geom_tile(color="black") +
    theme_classic()+
    theme(axis.line.y=element_blank(), axis.line.x=element_blank())+
    scale_fill_gradient2(low = "white", 
                         high = "navy", 
                         limits=c(0,0.20), 
                         oob = scales::squish) +
    guides(fill=guide_colourbar(title="Var"))+
       theme(text=element_text(family = "baskerville"), legend.position = "none")+
  scale_y_discrete(labels=c(1:30))+
ylab("")+
xlab("")+
  ggtitle("(b)")

#Change names of covariates
levels(fig1b$data$variable)<- c("Age", "Age(Onset)", "Sex", "Timepoint", "CTG")

#Merge figures
figure1<- ggarrange(fig1a, fig1b)
print(figure1)
```  

```{r}
#-----corr factors with clinical outcomes (spearman, BH) (Figure2)
#-----Generates two figures, uses the pval as mask for the heatmap with rho
#-----values
colorhue<- colorRampPalette(c("#440154FF" ,"white", "#FDE725FF"))(n=300)    

fig2a<- corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "r") +
      scale_x_discrete(labels=c(1:30)) +
      theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5))+
      scale_fill_gradient2(low="#440154FF", mid = "white", high = "#FDE725FF")+
      theme(text=element_text(family = "baskerville"))+
      guides(fill=guide_colourbar(title="rho"))+
      ggtitle("")
    
      
fig2b<- corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "pval") +
      scale_x_discrete(labels=c(1:30)) +
      theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5))+
      scale_fill_gradientn(colors=c("#440154FF","#FFFFFFFF"), values = c(0,0.05), na.value = "transparent")+
      theme(text=element_text(family = "baskerville"))+
      guides(fill=guide_colourbar(title="p-value"))+
      ggtitle("(b)")


      lev<- levels(fig2a$data$Var2)
      lev[grep("SMWT", lev)]<-"6MWT"
      levels(fig2a$data$Var2)<- lev

      lev<- levels(fig2b$data$Var2)
      lev[grep("SMWT", lev)]<-"6MWT"
      levels(fig2b$data$Var2)<- lev


      print(fig2a)
      print(fig2b) 

      fig2a$data$value[fig2b$data$value>0.05]<-0
      #Print Figure 2
      print(fig2a) 

```

```{r}
#-----Heatmap Factor 1 (rna) with Loess lm method V2Mode over samples 
tv<- plot_data_heatmap(MOFAobject_ran, 
                       view="rna", 
                       factor=1, 
                       features=30, 
                       cluster_cols= FALSE, 
                       #cluster_rows= FALSE, 
                  show_rownames= TRUE, show_colnames=TRUE, scale="row")
#tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)
 

clintoplot<-metadatadf[which(rownames(metadatadf)=="V2Mode"),]
rna_ss<- clintoplot[,colnames(clintoplot) %in% tv$gtable$grobs[[3]]$label]
rna_ss<- rna_ss[, tv$gtable$grobs[[3]]$label]
rna_ss<- t(rna_ss)
rna_ss<- as.data.frame(rna_ss)
rna_ss$rank<- c(1:nrow(rna_ss))
rna_ss$V2Mode<- as.numeric(rna_ss$V2Mode)

clinline<- ggplot(rna_ss, aes(x=rank, y=V2Mode))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(100,800))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
ylab("CTG")+ 
theme(axis.text.y = element_text(size=8), axis.text.x=element_blank(), panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), panel.border = element_rect( colour="grey", linewidth=1, fill=NA))

tv<- plot_data_heatmap2(MOFAobject_ran, 
                       view="rna", 
                       factor=1, 
                       features=30, 
                       cluster_cols= FALSE, 
                       cluster_rows= FALSE, 
                  show_rownames= TRUE, show_colnames=FALSE, scale="row", denoise=TRUE)
tv$gtable$grobs[[2]]$label<- mapRna(tv$gtable$grobs[[2]]$label)
dev.off()
grid::grid.draw(tv$gtable)


ggarrange(clinline+
              theme(plot.margin = margin(5,102,-15,5)), as.ggplot(gtable_add_cols(tv$gtable, unit(5,"null"),-1)), ncol = 1, nrow = 2, heights =c(0.10,0.9))


```


```{r}
#-----Heatmap Factor 23 (rna) with Loess lm method V2Mode over samples 
tv<- plot_data_heatmap(MOFAobject_ran, 
                       view="rna", 
                       factor=23, 
                       features=30, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=TRUE, scale="row")
tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)
 

clintoplot<-metadatadf[which(rownames(metadatadf)=="V2Mode"),]
rna_ss<- clintoplot[,colnames(clintoplot) %in% tv$gtable$grobs[[3]]$label]
rna_ss<- rna_ss[, tv$gtable$grobs[[3]]$label]
rna_ss<- t(rna_ss)
rna_ss<- as.data.frame(rna_ss)
rna_ss$rank<- c(1:nrow(rna_ss))
rna_ss$V2Mode<- as.numeric(rna_ss$V2Mode)

clinline<- ggplot(rna_ss, aes(x=rank, y=V2Mode))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(100,800))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
ylab("CTG")+
theme(axis.text.y = element_text(size=8), axis.text.x=element_blank(), panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), panel.border = element_rect( colour="grey", linewidth=1, fill=NA))

tv<- plot_data_heatmap(MOFAobject_ran,  
                       view="rna", 
                       factor=23, 
                       features=30, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=FALSE, scale="row", denoise=TRUE)
tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)
dev.off()
grid::grid.draw(tv$gtable)


ggarrange(clinline+
              theme(plot.margin = margin(5,102,-15,55)), as.ggplot(gtable_add_cols(tv$gtable, unit(5,"null"),-1)), ncol = 1, nrow = 2, heights =c(0.10,0.9))


```




```{r}
#-----Heatmap Factor 2 (rna) with Loess lm method 6MWT over samples 
tv<- plot_data_heatmap(MOFAobject_ran, 
                       view="rna", 
                       factor=2, 
                       features=25, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=TRUE, scale="row")
tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)


SMWT<-clinoutcomes_scaled[which(rownames(clinoutcomes_scaled)=="SMWT"),]
SMWT_rna_ss<- SMWT[,colnames(SMWT) %in% tv$gtable$grobs[[3]]$label]
SMWT_rna_ss<- SMWT_rna_ss[, tv$gtable$grobs[[3]]$label]
SMWT_rna_ss<- t(SMWT_rna_ss)
SMWT_rna_ss<- as.data.frame(SMWT_rna_ss)
SMWT_rna_ss$rank<- c(1:nrow(SMWT_rna_ss))

smwtline<-ggplot(SMWT_rna_ss, aes(x=rank, y=SMWT))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(-0.5,0.5))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
ylab("6MWT")+
theme(axis.text.y = element_text(size=8), axis.text.x=element_blank(), panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), panel.border = element_rect( colour="grey", linewidth=1, fill=NA))

tv<- plot_data_heatmap(MOFAobject_ran, 
                       view="rna", 
                       factor=2, 
                       features=25, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=FALSE, scale="row")
tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)
dev.off()
grid::grid.draw(tv$gtable)


ggarrange(smwtline+
              theme(plot.margin = margin(5,98,-15,55)), 
          as.ggplot(gtable_add_cols(tv$gtable, unit(5,"null"),-1)), 
          ncol = 1, nrow = 2, heights =c(0.10,0.9))


```

```{r}
#-----Heatmap Factor 3 (protein) with Loess lm method FDSS over samples 
tv<- plot_data_heatmap(MOFAobject_ran,  
                       view="protein", 
                       factor=3, 
                       features=30, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=TRUE, scale="row", denoise=TRUE)
tv$gtable$grobs[[3]]$label<- mapRna(tv$gtable$grobs[[3]]$label)


FDSS<-clinoutcomes_scaled[which(rownames(clinoutcomes_scaled)=="FDSS"),]
FDSS_rna_ss<- FDSS[,colnames(FDSS) %in% tv$gtable$grobs[[3]]$label]
FDSS_rna_ss<- FDSS_rna_ss[, tv$gtable$grobs[[3]]$label]
FDSS_rna_ss<- t(FDSS_rna_ss)
FDSS_rna_ss<- as.data.frame(FDSS_rna_ss)
FDSS_rna_ss$rank<- c(1:nrow(FDSS_rna_ss))

FDSSline<-ggplot(FDSS_rna_ss, aes(x=rank, y=FDSS))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(-2,2))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(
      colour = "grey",
      linewidth = 1,
      fill = NA
    )
  )

CISa<-clinoutcomes_scaled[which(rownames(clinoutcomes_scaled)=="CISactivity"),]
CISa_rna_ss<- CISa[,colnames(CISa) %in% tv$gtable$grobs[[3]]$label]
CISa_rna_ss<- CISa_rna_ss[, tv$gtable$grobs[[3]]$label]
CISa_rna_ss<- t(CISa_rna_ss)
CISa_rna_ss<- as.data.frame(CISa_rna_ss)
CISa_rna_ss$rank<- c(1:nrow(CISa_rna_ss))

CISaline<-ggplot(CISa_rna_ss, aes(x=rank, y=CISactivity))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(-2,2))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(
      colour = "grey",
      linewidth = 1,
      fill = NA
    )
  )

CISf<-clinoutcomes_scaled[which(rownames(clinoutcomes_scaled)=="CISFatigue"),]
CISf_rna_ss<- CISf[,colnames(CISf) %in% tv$gtable$grobs[[3]]$label]
CISf_rna_ss<- CISf_rna_ss[, tv$gtable$grobs[[3]]$label]
CISf_rna_ss<- t(CISf_rna_ss)
CISf_rna_ss<- as.data.frame(CISf_rna_ss)
CISf_rna_ss$rank<- c(1:nrow(CISf_rna_ss))

CISfline<-ggplot(CISf_rna_ss, aes(x=rank, y=CISFatigue))+
geom_smooth(method="loess", se=TRUE, color="347FC4")+
theme_minimal() +
coord_cartesian(ylim=c(-2,2))+
scale_x_discrete(lim=c(1:54))+
scale_y_continuous(position = "right")+
xlab("")+
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(
      colour = "grey",
      linewidth = 1,
      fill = NA
    )
  )



tv<- plot_data_heatmap(MOFAobject_ran, 
                       view="protein", 
                       factor=3, 
                       features=30, 
                       cluster_cols= FALSE, 
                  show_rownames= TRUE, show_colnames=FALSE, scale="row", denoise=TRUE)
tv$gtable$grobs[[3]]$label<- mapProtein(tv$gtable$grobs[[3]]$label)
dev.off()
grid::grid.draw(tv$gtable)


ggarrange(
  CISaline +
    theme(plot.margin = margin(5, 130, -15, 55)),
  as.ggplot(gtable_add_cols(tv$gtable, unit(5, "null"), -1)),
  CISfline +
    theme(plot.margin = margin(5, 130, -15, 55)),
  as.ggplot(gtable_add_cols(tv$gtable, unit(5, "null"), -1)),
  FDSSline +
    theme(plot.margin = margin(5, 130, -15, 55)),
  as.ggplot(gtable_add_cols(tv$gtable, unit(5, "null"), -1)),
  ncol = 1,
  nrow = 4,
  heights = c(0.10, 0.9)
)

 
```

```{r} 
#Table (features with highest weight) generation

for(i in c(1:30)){
print(plot_top_weights(MOFAobject_ran, view="phenotype", factor=i, nfeatures=22, scale=T) +
  theme(text=element_text(family = "baskerville"))) 

for(j in c("protein", "rna")){
print(plot_weights_mapped(MOFAobject_ran, view=j, factor=i, nfeatures=25, scale=T, genenames = genenames)+
  theme(text=element_text(family = "baskerville")))
}
}

```

```{r}
#Get all rna weights to perform GSEA Gprofiler (See Gprofiler GSEA script): 
#rnaweights<- get_weights(MOFAobject_ran, views="rna", factors="all")
#write.csv(rnaweights, file.choose())
```



```{r}
#-----Print all GSEA tables
for (i in as.character(c(1:30))){
  
  if(length(gsearesults[[i]])>0){
  p<-publish_gosttable(gsearesults[[i]])+
      ggtitle(paste0("Factor ", i))
  print(p)
  }
  
} 

```

```{r}
#Get list of top 25 highest weights for export to complete dataset (Model 1 script)

for(i in c(1:30)){
topweights_protein<- get_weights(MOFAobject_ran, 
                                 views="protein", 
                                 factors=i)
colnames(topweights_protein$protein)<- "factor"
topweights_protein<- as.data.frame(topweights_protein)
topweights_protein$protein<- rownames(topweights_protein)
topweights_protein$factor<- abs(topweights_protein$factor)
topweights_protein<- topweights_protein[order(topweights_protein$factor, decreasing = TRUE),]
protlist25<-paste(unlist(rownames(topweights_protein)[1:25]), collapse = ",")
print(paste0("Factor",i, ":", protlist25))
}
```




