---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
------
      title: "Multi-omics analysis"
      author: "Amir Chaman Baz"
      date: "4/5/2024"
      output:
      pdf_document: default
      html_document: default
------
#------------------------------------------------------------------------------------
# Script for Model 1; MOFA2 analysis of Protein+Phenomics using the complete data set
#------------------------------------------------------------------------------------
```{r}
#Select whether to run new MOFA2 analysis or load Mofaobject (Default=2)
cat("run new MOFA model(1) or load from file(2)?: ")
optionp<- readline()

if(!optionp==1){
  optionp<-2
}
```

#Load libraries
```{r}
      library(MOFA2)
      library(readxl)
      library(dplyr)
      library(purrr)
      library(ggplot2)
      library(reshape2)
      library(ggplotify)
      library(gridExtra)
      library(gplots)
      library(corrplot)
      library(MASS)
      library(VennDiagram)
      library(ggcorrplot)
      library(viridis)
      library(ggpmisc)
      library(sysfonts)
      library(showtext)
      library(ggpubr)
      library(psych)
      font.add("baskerville", regular="Libre_Baskerville(1)/LibreBaskerville-Regular.ttf", bold="Libre_Baskerville(1)/LibreBaskerville-Bold.ttf", italic = "Libre_Baskerville(1)/LibreBaskerville-Italic.ttf" )
      showtext.auto()
``` 

```{r}
#-----Import functions used for plot generation
source("importfunctions.R")

```

```{r}
#Load genenames database (EnsDb.Hsapiens.v86.sqlite)
library(RSQLite)
dbfile <- "EnsDb.Hsapiens.v86.sqlite"
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = dbfile)
ensembltables <- dbListTables(con)

ensembldf <- vector("list", length = length(ensembltables))

for (i in seq(along = ensembltables)) {
  ensembldf[[i]] <-
    dbGetQuery(
      conn = con,
      statement = paste("SELECT * FROM '", ensembltables[[i]], "'", sep = "")
    )
}

genenames <- data.frame(gene_id = ensembldf[[4]]$gene_id,
                        gene_name = ensembldf[[4]]$gene_name)
```

```{r}
#Load proteinnames from database (UniprotKB)
file<- "HUMAN_9606_idmapping_selected.tab/HUMAN_9606_idmapping_selected.tab"
proteindf<- read.csv(file, 
                    sep="\t", 
                    header=FALSE)
      
proteinnames<- data.frame(protein.id=proteindf$V1, 
                          protein.name=proteindf$V2)
      
file<- NULL

#Remove substring "_HUMAN" from proteinnames
proteinnames$protein.name<- substring(proteinnames$protein.name, 1, nchar(proteinnames$protein.name)-6)
```



```{r}
#-----Load data from external files
load(file.choose())
pdf$PatientID<- gsub("_V.", "", rownames(pdf))

load(file.choose())
rna_counts<- data.frame(v$E)

df<- read_excel(file.choose(), sheet=1)
df<- as.data.frame(df)
names(df)[names(df)=="PatientID...1"]<- "sample"
```


```{r}
#Calculate StroopInterference and TMT 
stroop2V2<- ((60-df$StroopCardIIErrorsV2) / 60) / df$StroopCardIITimeV2
stroop2V4<- ((60-df$StroopCardIIErrorsV4) / 60) / df$StroopCardIITimeV4
stroop3V2<- ((60-df$StroopCardIIIErrorsV2) / 60) / df$StroopCardIIITimeV2
stroop3V4<- ((60-df$StroopCardIIIErrorsV4) / 60) / df$StroopCardIIITimeV4
df$StroopInterferenceV2<- stroop3V2 / stroop2V2
df$StroopInterferenceV4<- stroop3V4 / stroop2V4

df$TMTV2<- df$TMTBV2 / df$TMTAV2
df$TMTV4<- df$TMTBV4 / df$TMTAV4


```


```{r}
#Selection of clinical outcomes from list
#seperate two timepoints and merge 

#select V2 vars only
clindf<- df[colnames(df)[grep("*\\V2$", colnames(df))]] 
#Select V4 vars only
clindfv4<- df[colnames(df)[grep("*\\V4$", colnames(df))]]


#remove all variables starting with Stroop
clindf<- clindf[-grep("^Stroop.*$", colnames(clindf))] 
clindfv4<- clindfv4[-grep("^Stroop.*$", colnames(clindfv4))] 
      
#Add Stroopinterference and TMT scores
clindf$StroopInterferenceV2<- df$StroopInterferenceV2
clindf$TMTV2<- df$TMTV2

clindfv4$StroopInterferenceV4<- df$StroopInterferenceV4
clindfv4$TMTV4<- df$TMTV4
      
#Remove TMTA en TMTB, PreBorg, MIRS, MCgillpain, l5ENMO, M5ENMO, SIPScore
omitlist<- c(paste0("TMTBV2|TMTAV2|PreBORGV2|MIRSV2|McGillPainV2|",
                          "L5ENMOV2|M5ENMOV2|SIPScoreV2"))
clindf<- clindf[!grepl(omitlist, colnames(clindf))]
      
omitlist<- c(paste0("TMTBV4|TMTAV4|PreBORGV4|MIRSV4|McGillPainV4|",
                          "L5ENMOV4|M5ENMOV4|SIPScoreV4"))
clindfv4<- clindfv4[!grepl(omitlist, colnames(clindfv4))]

#Add samplenames to both df      
samplenames<-substring(df$sample, 1, nchar(df$sample)-1)
rownames(clindf)<-paste0(samplenames,"_V2")
rownames(clindfv4)<-paste0(samplenames,"_V4")
      
```

```{r}
#-----Remove V2/V4 from colnames, merge df's
colnames(clindfv4)<- substring(colnames(clindfv4),
                                     1, nchar(colnames(clindfv4))-2)
colnames(clindf)<- substring(colnames(clindf), 
                                     1, nchar(colnames(clindf))-2)
clinicaloutcomescoresdf<- rbind(clindf, clindfv4)
```



```{r}
#Filtered phenotypes (Apply IQR_filter to clinicaloutcomes)
clinfiltered<- lapply(clinicaloutcomescoresdf, function(x)IQR_filter(x,3)) 
clinfiltered<- as.data.frame(clinfiltered)
rownames(clinfiltered)<- rownames(clinicaloutcomescoresdf)

numfiltered<-sum(is.na(clinfiltered)) - sum(is.na(clinicaloutcomescoresdf))
paste0(numfiltered, " outliers filtered ") #Should be 5 outliers

clinicaloutcomescoresdf<- clinfiltered


```

```{r}
#Make new dataframe clinoutcomes with transformed variables
clinoutcomes<-clinicaloutcomescoresdf

clinoutcomes$CISFatigue<- clinicaloutcomescoresdf$CISFatigue^2 
clinoutcomes$BDIFs<- sqrt(clinicaloutcomescoresdf$BDIFs)
clinoutcomes$SSLNScore<- 1/sqrt(clinicaloutcomescoresdf$SSLNScore)
clinoutcomes$SSLDScore<- 1/sqrt(clinicaloutcomescoresdf$SSLDScore)
clinoutcomes$JFCS<- 1/sqrt(clinicaloutcomescoresdf$JFCS)
clinoutcomes$CISactivity<- clinicaloutcomescoresdf$CISactivity^2


```


```{r}
#-----Z-score scaling for each clinical outcome measurement by variable
clinoutcomes_scaled<- as.data.frame(lapply(clinoutcomes, 
                                                 function(x) 
                                                          scale(x, 
                                                          center=TRUE, 
                                                          scale = TRUE)))
rownames(clinoutcomes_scaled)<- rownames(clinoutcomes)


```

```{r}
#Z-score scaling for each rpc protein 
rpc_scaled<- as.data.frame(apply(rpc, 1, 
                                  function(x) 
                                      scale(x, 
                                            center=TRUE, 
                                            scale = TRUE)))


rpc_scaled<-t(rpc_scaled)
colnames(rpc_scaled)<- colnames(rpc)
rpc_scaled<- as.data.frame(rpc_scaled)
```

```{r}
#Inverse clinical outcomes so they have have the same directionality (higher is better)
clinoutcomes_scaled$MDHI <- clinoutcomes_scaled$MDHI*(-1)
clinoutcomes_scaled$CISFatigue <- clinoutcomes_scaled$CISFatigue*(-1)
clinoutcomes_scaled$INQOLQolScore <- clinoutcomes_scaled$INQOLQolScore*(-1)
clinoutcomes_scaled$BDIFs <- clinoutcomes_scaled$BDIFs*(-1)
clinoutcomes_scaled$AEScScore <- clinoutcomes_scaled$AEScScore*(-1)
clinoutcomes_scaled$TMT <- clinoutcomes_scaled$TMT*(-1)
clinoutcomes_scaled$ASBQ <- clinoutcomes_scaled$ASBQ*(-1)
clinoutcomes_scaled$SSLDScore <- clinoutcomes_scaled$SSLDScore*(-1)
clinoutcomes_scaled$SSLNScore <- clinoutcomes_scaled$SSLNScore*(-1)
clinoutcomes_scaled$JFCS <- clinoutcomes_scaled$JFCS*(-1)
clinoutcomes_scaled$IMQ <- clinoutcomes_scaled$IMQ*(-1)
clinoutcomes_scaled$CISactivity <- clinoutcomes_scaled$CISactivity*(-1)
clinoutcomes_scaled$FDSS <- clinoutcomes_scaled$FDSS*(-1)

#CSI and AESI lower is better, added this
clinoutcomes_scaled$CSI <- clinoutcomes_scaled$CSI*(-1)
clinoutcomes_scaled$AESI <- clinoutcomes_scaled$AESI*(-1)

```

```{r}
#-----Create backups of rna and rpc 
rna_counts2<- rna_counts
rpc2<-rpc_scaled

      
clinoutcomes_scaled<-t(clinoutcomes_scaled)
clinoutcomes_scaled<- as.data.frame(clinoutcomes_scaled)
      
rna_counts3<- rna_counts2
clinoutcomes_scaled3<- clinoutcomes_scaled
rpc3<-rpc2
     
      
```

```{r}
#Make rpc and rna_counts have equal samples by adding NA columns for the 
#missing samples in each
      
mask<- !(colnames(clinoutcomes_scaled) %in% colnames(rpc2))
missingsamples<- colnames(clinoutcomes_scaled)[mask]
      for (y in missingsamples ){
            rpc2[[y]]<- NA
          }
      
      mask<- !(colnames(clinoutcomes_scaled) %in% colnames(rna_counts2))
      missingsamples<- colnames(clinoutcomes_scaled)[mask]
      for (y in missingsamples ){
            rna_counts2[[y]]<- NA
          }
      
      rna_counts2<- rna_counts2[,order(colnames(rna_counts2))]
      rpc2<- rpc2[,order(colnames(rpc2))]
      clinoutcomes_scaled<- clinoutcomes_scaled[, order(colnames(clinoutcomes_scaled))]
```
  
```{r}
#Build MOFA Object and Set Options (Slow convergence, max iteration 10000)
      mofadata<-list(#rna = as.matrix(rna_counts2), 
                     protein= as.matrix(rpc2), 
                     phenotype=as.matrix(clinoutcomes_scaled))
      MOFAobject<- create_mofa(mofadata)
      
      
      data_opts<- get_default_data_options(MOFAobject)
      model_opts<- get_default_model_options(MOFAobject)
      model_opts$num_factors<- 15 #Number of factors
      train_opts<- get_default_training_options(MOFAobject)
      train_opts$convergence_mode <- "slow"
      train_opts$maxiter<- 10000
      
      MOFAobject <- prepare_mofa(MOFAobject, 
                                 data_options= data_opts, 
                                 model_options = model_opts, 
                                 training_options = train_opts)


```
  


```{r}
#load model:
mofaobjectfile<-file.choose()
      if(optionp==2){
      MOFAobject_ran<- load_model(mofaobjectfile)}else if(optionp==1){
        MOFAobject_ran<- run_mofa(MOFAobject, 
                           outfile=file.choose(), 
                           use_basilisk = FALSE)
                          }
```
#-------------------------------------------------------------------------------


```{r}
#load metadata in MOFAobject_ran: 
      df2<-df
      df2$sample<-paste0(substring(df2$sample, 1, nchar(df2$sample)-1))


      df3<-df2[1,]
      df3$timepoint<- ""
      df3<-df3[0,]

#For each sample V2 and V4, double metadata so every sample has 
#its own metadata entry
      for(g in df2$sample){
          df3<-rbind(df3, 
                     cbind(sample=paste0(g, "_V2"), df2[which(df2$sample==g),-1], timepoint=0), 
                     cbind(sample=paste0(g, "_V4"), df2[which(df2$sample==g),-1], timepoint=1))
          }


      df3$CTGDiagnostic<- as.numeric(df3$CTGDiagnostic)
      df3$LogCTG_s<- scale(log(df3$CTGDiagnostic), center=TRUE, scale=TRUE)
      df3$AgeAtOnset<- as.numeric(df3$AgeAtOnset)
      metadatadf<- cbind(df3[!colnames(df3) %in% rownames(clinoutcomes_scaled)], t(clinoutcomes_scaled))
      samples_metadata(MOFAobject_ran) <- metadatadf
      metadatadf<- t(metadatadf)
      metadatadf<- as.data.frame(metadatadf) 
      colnames(metadatadf)<- metadatadf[which(rownames(metadatadf)=="sample"),] 
```
 
#-------------------------------------------------------------------------------
#-----Start of Plot generation

```{r}
#manual variance data:
r2 <- calculate_variance_explained(MOFAobject_ran)
r2perfactor<-as.matrix(r2$r2_per_factor$group1)
r2perfactor<- melt(r2perfactor)
```

```{r}
#LM function factors ~ covariates
#returns both the R2 as well as the BH adjusted p.val
listn <-Corr_factors_R2(MOFAobject_ran, 
                covariates = c(#"TreatmentCode", 
                               "V2Age",
                               "AgeAtOnset",
                               "SexCode", 
                            "timepoint",
                               "CTGDiagnostic"), 
                plot = "r")
Z<- listn[[1]] #Retrieve Factor values per sample
covariates <- listn[[2]] #Get covariates
Z<-as.data.frame(Z)


lmheatmap<-covariates[0,] #Empty df with Covariate names as colnames
lmheatmap_pval<- covariates[0,]

for(j in 1:15){
    l<-list()
    pval<- list()
    for(i in colnames(lmheatmap))
    {
        
        testdf<- data.frame(Z=Z[[j]], cov=covariates[[i]])
        
        lmmodel<- summary(lm(cov ~ Z, data = testdf))
        l<- c(l, lmmodel$r.squared)
        pval<- c(pval, lmmodel$coefficients[2,4])
    }
    
    l<- as.data.frame(t(l))
    colnames(l)<- colnames(lmheatmap)
    lmheatmap<- rbind(lmheatmap, l)
    
    pval<- as.data.frame(t(pval))
    colnames(pval)<- colnames(lmheatmap_pval)
    lmheatmap_pval <- rbind(lmheatmap_pval, pval)
    
}
lmheatmap<- as.data.frame(lmheatmap)
#lmheatmap$TreatmentCode<- as.numeric(lmheatmap$TreatmentCode)
lmheatmap$V2Age<- as.numeric(lmheatmap$V2Age)
lmheatmap$AgeAtOnset<- as.numeric(lmheatmap$AgeAtOnset)
lmheatmap$SexCode<- as.numeric(lmheatmap$SexCode)
lmheatmap$timepoint<- as.numeric(lmheatmap$timepoint)
lmheatmap$CTGDiagnostic<- as.numeric(lmheatmap$CTGDiagnostic)

rownames(lmheatmap)<-r2perfactor$Var1[1:15]
lmheatmap$factor<-r2perfactor$Var1[1:15]
lmmelt<- melt(lmheatmap, id.vars = "factor")

lmheatmap_pval<- as.data.frame(lmheatmap_pval)
lmheatmap_pval$V2Age<- as.numeric(lmheatmap_pval$V2Age)
lmheatmap_pval$AgeAtOnset<- as.numeric(lmheatmap_pval$AgeAtOnset)
lmheatmap_pval$SexCode<- as.numeric(lmheatmap_pval$SexCode)
lmheatmap_pval$timepoint<- as.numeric(lmheatmap_pval$timepoint)
lmheatmap_pval$CTGDiagnostic<- as.numeric(lmheatmap_pval$CTGDiagnostic)

rownames(lmheatmap_pval)<-r2perfactor$Var1[1:15]
lmheatmap_pval$factor<-r2perfactor$Var1[1:15]

lmheatmap_pval<- melt(lmheatmap_pval, value.name = "pval")
lmheatmap_pval<- as.data.frame(lmheatmap_pval)
lmheatmap_pval$pval_adjusted<-p.adjust(lmheatmap_pval$pval, method="BH")
```

```{r}
#get weights protein in Factor 3
#Load DvA single omic data (6MWT lead hits) for rank plots

topweights_protein_f3<- get_weights(MOFAobject_ran, views="protein", factors=3)
topweights_protein_f3<- as.data.frame(topweights_protein_f3)
topweights_protein_f3$protein<- rownames(topweights_protein_f3)
topweights_protein_f3$Factor3<- abs(topweights_protein_f3$Factor3)
topweights_protein_f3<- topweights_protein_f3[order(abs(topweights_protein_f3$Factor3), decreasing = TRUE),]


dva_hits_SMWT<- read.csv(file.choose())

topweights_protein_f3$rank<- 1:nrow(topweights_protein_f3)

hitlist<- topweights_protein_f3$protein[1:73][topweights_protein_f3$protein[1:73] %in% dva_hits_SMWT$x]

#print(dva_hits_SMWT[dva_hits_SMWT$x %in% hitlist,]) 

#venn.diagram(x= list(topweights_protein_f3$protein[1:73], dva_hits_SMWT$x), filename="052724_SMWT_f3_venn.png", category.names = c("MOFA2", "DvA hits"), fill=c(ggplot2::alpha("#440154FF", 0.3), ggplot2::alpha("#FDE725FF", 0.3)), output=TRUE)

```


```{r}
#get weights protein in Factor 2 and 4
#Load DvA single omic data (CTG lead hits) for rank plots
topweights_protein_f4<- get_weights(MOFAobject_ran, views="protein", factors=4)
topweights_protein_f4<- as.data.frame(topweights_protein_f4)
topweights_protein_f4$protein<- rownames(topweights_protein_f4)
topweights_protein_f4$Factor4<- abs(topweights_protein_f4$Factor4)
topweights_protein_f4<- topweights_protein_f4[order(topweights_protein_f4$Factor4, decreasing = TRUE),]

topweights_protein_f2<- get_weights(MOFAobject_ran, views="protein", factors=2)
topweights_protein_f2<- as.data.frame(topweights_protein_f2)
topweights_protein_f2$protein<- rownames(topweights_protein_f2)
topweights_protein_f2$Factor4<- abs(topweights_protein_f2$Factor2)
topweights_protein_f2<- topweights_protein_f2[order(topweights_protein_f2$Factor2, decreasing = TRUE),]



dva_hits_CTG<- read.csv(file.choose())

topweights_protein_f4$rank<- 1:nrow(topweights_protein_f4)

#What overlap topweights ~ CTG hits DvA in top 50
#print(topweights_protein_f4[1:50,][topweights_protein_f4$protein[1:50] %in% dva_hits_CTG$x[1:50],])


 

  
``` 

```{r}
#-----Get all weights (protein, phenotype)
weights<-get_weights(MOFAobject_ran, views="all", factors="all")
protein.weights<-as.data.frame(weights$protein)
#rna.weights<-as.data.frame(weights$rna)
phenotype.weights<-as.data.frame(weights$phenotype)
```



#-------------------------------------------------------------------------------

```{r}
#-----Sanity check; verify whether factors are largely uncorrelated:
      plot_factor_cor(MOFAobject_ran)
```

*Usage of function "regressor()":*
soutcome= String, Outcome or covariate to correlate from MOFA2 model metadata
nfactor= Integer, Factor to use to retrieve top features
nfeature=Either an integer depicting the top n features to extract from factor, or a (list of) string of specific features to check
colby= String, covariate to color the points
returncorr= Return spearman correlations, not figures. Default=FALSE

```{r}
#Numbers used in paper; CTG~factor 2 (protein only)(Model1)
regressor("V2Mode", nfactor = 2, nfeature = 9, colby="timepoint", returncorr=TRUE)
regressor("V2Mode", nfactor = 2, nfeature = 9, colby="timepoint", returncorr=TRUE)$p.adj 
```

```{r}
#Numbers used in paper; 6MWT~factor 3 (protein only)(Model 1)
regressor("SMWT", nfactor = 3, nfeature = 9, colby="timepoint", returncorr=TRUE)
regressor("SMWT", nfactor = 3, nfeature = 9, colby="timepoint", returncorr=TRUE)$p.adj
```


#Figures:
#Protein variance over top 5 29.3%
#phenotype variance 33.9%

```{r}
#variance per factor (Model 1) (Figure 1 appendix chapter)
fig1a<-plot_variance_explained(MOFAobject_ran, plot_total = F)+
  theme(axis.text.x = element_text(size=12), axis.text.y = element_text(size=12),text=element_text(family = "baskerville"))+
  ggtitle("(a)")

fig1b  <-ggplot(lmmelt, 
       aes(x = variable, y = factor, fill = value)) +
    geom_tile(color="black") +
    theme_classic()+
    theme(axis.line.y=element_blank(), axis.line.x=element_blank())+
    scale_fill_gradient2(low = "white", 
                         high = "navy", 
                         limits=c(0,0.15), 
                         oob = scales::squish) +
    guides(fill=guide_colourbar(title="R2"))+
       theme(axis.text.x = element_text(size=12), axis.text.y = element_text(size=12),text=element_text(family = "baskerville"), legend.position = "none")+
ylab("")+
xlab("")+
  ggtitle("(b)")
levels(fig1b$data$variable)<- c("Age", "Age(Onset)", "Sex", "Timepoint", "CTG") #Change column names

figure1<-ggarrange(fig1a, fig1b)
print(figure1)
```

  
```{r}
#Spearman correlations of factors with clinical outcomes (Figure 2)
#Mask spearman rho by p<0.05 and merge figures 
colorhue<- colorRampPalette(c("#440154FF" ,"white", "#FDE725FF"))(n=300)    
		  
#corr factors with clinical outcomes
fig2a<- corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "r") +
    scale_x_discrete(labels=c(1:15)) +
    theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5))+
    scale_fill_gradient2(low="#440154FF", mid = "white", high = "#FDE725FF")+
    theme(text=element_text(family = "baskerville"))+
    guides(fill=guide_colourbar(title="rho"))+
    ggtitle("(a)")
    
fig2b<- corr_factors_cov_spearman(MOFAobject_ran, covariates = rownames(clinoutcomes_scaled), plot = "p") +
    scale_x_discrete(labels=c(1:15)) +
    theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5))+
    scale_fill_gradientn(colors=c("#440154FF","#FFFFFFFF"), values = c(0,0.05), na.value = "transparent")+
    theme(text=element_text(family = "baskerville"))+
     guides(fill=guide_colourbar(title="p-value"))+
    ggtitle("(b)")
      
lev<- levels(fig2a$data$Var2)
lev[2]<-"6MWT"
levels(fig2a$data$Var2)<- lev

lev<- levels(fig2b$data$Var2)
lev[2]<-"6MWT"
levels(fig2b$data$Var2)<- lev

#Heatmap weights per factor
#fig2b<- heatmap.2(weights$phenotype, Colv = FALSE, col=colorhue, tracecol = '#347FC4', sepcolor  = "black", colsep=1:ncol(weights$phenotype), rowsep=1:nrow(weights$phenotype), main="(b)", key.title = "Weight" ) 

print(fig2a)
print(fig2b) 


ggarrange(fig2a, fig2b, ncol=2, nrow=1)

fig2a$data$value[fig2b$data$value>0.05]<-0

print(fig2a)


#print(spearman_pval)
```
 
```{r}
#Weight tables for factor 2 (model 1)(merged)
tablef2_1<-plot_weights_mapped(MOFAobject_ran, view="protein", factor=2, nfeatures=25, scale=T)+
  theme(text=element_text(family = "baskerville"))
tablef2_2<- plot_top_weights(MOFAobject_ran, view="phenotype", factor=2, nfeatures=22, scale=T) +
  theme(text=element_text(family = "baskerville"))  

lv<- levels(tablef2_2$data$feature_id)
lv[grep("SMWT", lv)]<- "6MWT"
levels(tablef2_2$data$feature_id)<- lv

ggarrange(tablef2_1, tablef2_2, ncol=2, nrow=1)

#Correlation plots for factor 2 (model 1)
regressor("V2Mode", nfactor = 2, nfeature = 9, colby="timepoint", ncoln = 3, nrown = 3) 

```  
 
```{r}
#Weight tables for factor 2 (model 1)(merged)
tablef3_1<-plot_weights_mapped(MOFAobject_ran, view="protein", factor=3, nfeatures=25, scale=T)+
  theme(text=element_text(family = "baskerville"))
tablef3_2<- plot_top_weights(MOFAobject_ran, view="phenotype", factor=3, nfeatures=22, scale=T) +
  theme(text=element_text(family = "baskerville"))  

lv<- levels(tablef3_2$data$feature_id)
lv[grep("SMWT", lv)]<- "6MWT"
levels(tablef3_2$data$feature_id)<- lv


ggarrange(tablef3_1, tablef3_2, ncol=2, nrow=1)

#Correlation plots for factor 2 (model 1)
regressor("SMWT", nfactor = 3, nfeature = 9, colby="timepoint", ncoln = 3, nrown = 3)  


``` 

```{r}
#rankplots for validating model 
print(rankplot(mapProtein(dva_hits_CTG$x[1:40]), mapProtein(topweights_protein_f2$protein[1:40])))
print(rankplot(mapProtein(dva_hits_CTG$x[1:40]), mapProtein(topweights_protein_f4$protein[1:40])))

print(rankplot(mapProtein(dva_hits_SMWT$x[1:40]), mapProtein(topweights_protein_f3$protein[1:40])))
```


